---
title: "PS_ Probability and Hypothesis Testing for TAs"
author: "Pablo E. Gutiérrez-Fonseca"
date: "2023-08-12"
output: pdf_document
---

# **R practice.<br />**
\vspace{1em}

**Load the water pollution data into R.<br />** 
```{r, echo=FALSE}
df_mammals <- read.csv('C:/Users/Guti/OneDrive - University of Vermont/Curriculum/07_ Cursos/NR-2400/Problem set_ M5 Probability and Hypothesis Testing/Sleeping Mammals.csv')
summary(df_mammals)
```
\newpage

1. Based on the summary results in JMP, mammals spend most of their time in which type of
sleep phase (dreaming or non-dreaming)?  
```{r}
mean(df_mammals$NonDreaming, na.rm = TRUE)
mean(df_mammals$Dreaming, na.rm = TRUE)
```
\vspace{1em}

2. Which type of sleep phase has the highest variability across the species included here
(dreaming or non-dreaming)?  
Non-dreaming  
```{r}
sd(df_mammals$NonDreaming, na.rm = TRUE)
sd(df_mammals$Dreaming, na.rm = TRUE)
```
\vspace{1em}


3. Enter the p-value for the goodness of fit test for the **Dreaming variable** (NOTE….just enter the number (no letters, symbols, equal signs etc)…also be careful of decimal places.
```{r}
shapiro.test(df_mammals$Dreaming)
```
\vspace{1em}

4. Based on this goodness of fit test, is the **dreaming variable** normally distributed?
**No**
\vspace{1em}

5. Enter the new p-value for the goodness of fit test for the **Dreaming variable** when the outliers are excluded (NOTE….just enter the number (no letters, symbols, equal signs
etc)…also be careful of decimal places.
```{r}
quartiles <- quantile(na.omit(df_mammals$Dreaming), probs = c(0.25, 0.75))
IQR <- IQR(na.omit(df_mammals$Dreaming))
Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 

new_Dreaming <- subset(df_mammals$Dreaming, df_mammals$Dreaming > Lower & df_mammals$Dreaming < Upper)

shapiro.test(new_Dreaming)
```

