---
title: "PS_Indpendent and Dependent t-test-for-TAs"
author: "Pablo E. Gutiérrez-Fonseca"
date: "2023-08-11"
output: pdf_document
---

**R practice.<br />**  
<br />  
<br /> 


**Install packages.<br />**  
```{r}
#install.packages("car")
library(car)
library(ggplot2)
```

<br />
<br />
<br />

**Load the water pollution data into R.<br />**   
```{r, echo=FALSE}
df_bird <- read.csv('C:/Users/Guti/OneDrive - University of Vermont/Curriculum/07_ Cursos/NR-2400/Problem set_ Independent and Dependent t-tests/Bird_Migration.csv')
head(df_bird)
```
<br />
<br />
<br />

**Analyze data distribution. Is your pre-weight data normally distributed?**  
```{r}
shapiro.test(df_bird$Pre_weight)
```
<br />
<br />
<br />

**Test to see if you meet the assumption of equal variance for this test. Based on the results of this test for unequal variance, which test should you use?**  
```{r}
# Perform Levene's test
result <- leveneTest(Pre_weight~Gender, data=df_bird)
print(result)
```
<br />
<br />
<br />  

```{r}
p_value <- result$Pr[1]
# Significance level
alpha <- 0.05

# Interpret the result
if (p_value < alpha) {
  cat("Reject null hypothesis: Variances are significantly different.\n")
} else {
  cat("Fail to reject null hypothesis: Variances are not significantly different.\n")
}
```
<br />
<br />
<br />  

**Run this analysis. What is the test statistic (obtained) value for the test you selected?**<br />  
 Use alternative to determine your hypothesis. Default= "two.sided".<br />   
 alternative = c("two.sided", "less", "greater"),
```{r}
# Perform t-test
t_test_result <- t.test(Pre_weight~Gender, data=df_bird,
                         alternative = "less")
print(t_test_result)
```

```{r}
# Create a boxplot plot using ggplot2
ggplot(df_bird, aes(x = Gender, y = Pre_weight)) +
  geom_boxplot()  # Add jitter for better visualization
```

