---
title: "PS: Indpendent and Dependent t-test for TAs"
author: "Pablo E. Gutiérrez-Fonseca"
date: "2023-08-11"
output: pdf_document
---

# **R practice.<br />**  
\vspace{1em}

**Install packages.<br />**  
```{r}
library(car)
library(ggplot2)
library(dplyr)
```
\vspace{1em}

**Load the Bird Migration data into R.<br />**   
```{r, echo=FALSE}
df_bird_all <- read.csv('C:/Users/Guti/OneDrive - University of Vermont/Curriculum/07_ Cursos/NR-2400/Problem set_ Independent and Dependent t-tests/Bird_Migration.csv')
head(df_bird_all)
```
\vspace{1em}

Make a subgroup of Weight before migration (Pre_weight).
```{r}
df_bird_Pre_weight <- df_bird_all %>% 
  filter(Treatment == 'Pre_weight')

summary(df_bird_Pre_weight)
```
\newpage



# **Part I.**<br />  
**2. Analyze data distribution. Is your pre-weight data normally distributed?**  
```{r}
shapiro.test(df_bird_Pre_weight$value)
```
\vspace{1em}

**3. Test to see if you meet the assumption of equal variance for this test. Based on the results of this test for unequal variance, which test should you use?**  
```{r}
# Perform Levene's test
result <- leveneTest(value~Gender, data=df_bird_Pre_weight)
print(result)
```
\vspace{1em} 

```{r}
p_value <- result$Pr[1]
# Significance level
alpha <- 0.05

# Interpret the result
if (p_value < alpha) {
  cat("Reject null hypothesis: Variances are significantly different.\n")
} else {
  cat("Fail to reject null hypothesis: Variances are not significantly different.\n")
}
```
\vspace{1em}

**4. Run this analysis. What is the test statistic (obtained) value for the test you selected?**<br />  
 Use alternative to determine your hypothesis. Default= "two.sided".<br />   
 alternative = c("two.sided", "less", "greater"),
```{r}
# Perform t-test
t_test_result <- t.test(value~Gender, alternative = "two.sided",
                         var.equal=TRUE, data=df_bird_Pre_weight)
print(t_test_result)
```
\vspace{1em}
We hypothesized that there is a difference in the pre-weights of male and female songbirds. However, an independent two-tailed pooled t-test indicated that there was not a significant difference between genders (t (26) = -0.78654, p= 0.4387). It is possible that low power limited the ability of this test to detect significant differences. 
\vspace{1em}

```{r}
# Create a boxplot plot using ggplot2
ggplot(df_bird_Pre_weight, aes(x = Gender, y = value)) +
  geom_boxplot()  
```
\newpage




# **Part II.**<br />  
**Analyze data distribution. Is your pre-weight data normally distributed?**  
```{r}
shapiro.test(df_bird_Pre_weight$value)
```
\vspace{1em}

**Test to see if you meet the assumption of equal variance for this test. Based on the results of this test for unequal variance, which test should you use?**  
```{r}
# Perform Levene's test
result <- leveneTest(value~Species, data=df_bird_Pre_weight)
print(result)
```
\vspace{1em} 

```{r}
p_value <- result$Pr[1]
# Significance level
alpha <- 0.05

# Interpret the result
if (p_value < alpha) {
  cat("Reject null hypothesis: Variances are significantly different.\n")
} else {
  cat("Fail to reject null hypothesis: Variances are not significantly different.\n")
}
```
\vspace{1em}

**8. Run this analysis. What is the test statistic (obtained) value for the test you selected?**<br />  
 Use alternative to determine your hypothesis. Default= "two.sided".<br />   
 alternative = c("two.sided", "less", "greater"),
```{r}
# Perform t-test
t_test_result <- t.test(value~Species, alternative = "two.sided",
                         var.equal=TRUE, data=df_bird_Pre_weight)
print(t_test_result)
```
\vspace{1em}
We hypothesized that there is a difference in the pre-weights of Hylocichla and Progne species. An independent two-tailed pooled t-test indicates that Hylocichla does have significantly lower pre-weights than Progne individuals (t(26) = -2.891043, p = 0.007655).  This difference is also meaningful (Cohen’s D effect size = 0.96) This indicates that any analyses of weight changes during migration may want to consider which species is examined based on their inherent difference in size.  
\vspace{1em}

```{r}
# Create a boxplot plot using ggplot2
ggplot(df_bird_Pre_weight, aes(x = Species, y = value)) +
  geom_boxplot()  
```
\newpage


# **Part III.<br />**  
**12. Analyze data distribution. Is your pre-weight data normally distributed?**  
```{r}
shapiro.test(df_bird_all$value)
```
\vspace{1em}


**Test to see if you meet the assumption of equal variance for this test. Based on the results of this test for unequal variance, which test should you use?**  
```{r}
# Perform Levene's test
result_paired_t_test <- leveneTest(value~Treatment, data=df_bird_all)
print(result_paired_t_test)
```
\vspace{1em} 

```{r}
p_value <- result_paired_t_test$Pr[1]
# Significance level
alpha <- 0.05

# Interpret the result
if (p_value < alpha) {
  cat("Reject null hypothesis: Variances are significantly different.\n")
} else {
  cat("Fail to reject null hypothesis: Variances are not significantly different.\n")
}
```
\vspace{1em}

**Perform and Print the paired t-test result.**
```{r}
t_paired_result <- t.test(value ~ Treatment, paired = TRUE, 
                          data = df_bird_all)
print(t_paired_result)
```
\vspace{1em}
We hypothesized that songbirds lose body mass during migration (quantified as a comparison of pre-and post-weights). A dependent one-tailed t-test indicated this was true as there was a significant loss of body mass for songbirds during migration (t (27) = -3.4184, p =0.002013).  This difference is also moderately meaningful (Cohen’s D effect size = 0.678).
\vspace{1em}


```{r}
# Create a boxplot plot using ggplot2
ggplot(df_bird_all, aes(x = Treatment, y = value)) +
  geom_boxplot()  
```
\newpage

# **Part IV.**  
**From the data.frame make two groups (females and males)**
```{r}
df_bird_female <- df_bird_all %>% 
  filter(Gender == 'F')
```
\vspace{1em}

```{r}
df_bird_male <- df_bird_all %>% 
  filter(Gender == 'M')
```
\vspace{1em}

**Assumptions for Females**
```{r}
shapiro.test(df_bird_female$value)
leveneTest(value~Treatment, data=df_bird_female)
```
\vspace{1em}

**Perform and Print the paired t-test result**
```{r}
t_paired_female <- t.test(value ~ Treatment, paired = TRUE, 
                          alternative= 'less', 
                          data = df_bird_female)
print(t_paired_female)
```
\vspace{2em}

**Assumptions for Males**
```{r}
shapiro.test(df_bird_male$value)
leveneTest(value~Treatment, data=df_bird_male)
```
\vspace{1em}

**Perform and Print the paired t-test result**
```{r}
t_paired_male <- t.test(value ~ Treatment, paired = TRUE, 
                        alternative= 'less',
                        data = df_bird_male)
print(t_paired_male)
```
\vspace{1em}
We wanted to see if males and females lose body mass during migration (measured as change in body weight).  Two separate one-tailed, dependent t-tests indicate that  both males (t (13) = -3.16013, p = 0.003762) and females (t (13) = -1.9868, p=0.03422) lose body weight during migration, although this loss if more significant in males.  This difference, while highly meaningful for males, was not meaningful for females (Cohen’s D effect size = 1.02  and 0.25   respectively for males and females).  This indicates that while both genders lose weight, females may spend more time foraging along the route to maintain their weight for breeding.


