---
title: "PS_ Descriptive Statistics for TAs"
author: "Pablo E. Gutiérrez-Fonseca"
date: "2023-08-13"
output: pdf_document
---




# **R practice.<br />**
\vspace{1em}

**Install packages.<br />**  
```{r}
library(tidyverse)
library(lessR)
library(DT)
library(e1071)
```
\vspace{1em}

**Load the water pollution data into R.<br />** 
```{r, echo=FALSE}
df_snow <- read.csv('Snow Depth data.csv')
summary(df_snow)
```
\newpage

1. the **MEAN Yearly Mean Snow Depth**.  
```{r}
round(mean(df_snow$Mean_Dec_May_snow_depth),2)
```
\vspace{1em}

1.1. the **MEDIAN Yearly Mean Snow Depth**.  
```{r}
round(median(df_snow$Mean_Dec_May_snow_depth),2)
```
\vspace{1em}


2. Still in Excel, now calculate the **STANDARD DEVIATION** for the Yearly Mean Snow
Depth.  
```{r}
round(sd(df_snow$Mean_Dec_May_snow_depth),2)
```
\vspace{1em}

2.1
```{r}
round(var(df_snow$Mean_Dec_May_snow_depth),2)
```


3. What is the **INTER-QUARTILE RANGE** for **Yearly Mean snow Depth**?
```{r}
# Calculate the IQR for the specified column
round(iqr_value <- IQR(df_snow$Mean_Dec_May_snow_depth),2)

# Print the IQR
print(iqr_value)
```
\vspace{1em}

4. Using the Interquartile range technique, how many **OUTLIER** years are there in your
**Yearly Mean Snow Depth Data**?   
```{r}
# Calculate the quartiles and IQR
q1 <- quantile(df_snow$Mean_Dec_May_snow_depth, 0.25, na.rm = TRUE)
q3 <- quantile(df_snow$Mean_Dec_May_snow_depth, 0.75, na.rm = TRUE)
iqr <- q3 - q1

# Define the lower and upper bounds for outliers
lower_bound <- q1 - 1.5 * iqr
upper_bound <- q3 + 1.5 * iqr

# Identify outlier years
outlier_years <- df_snow$Mean_Dec_May_snow_depth < lower_bound | df_snow$Mean_Dec_May_snow_depth > upper_bound

# Count the number of outlier years
num_outliers <- sum(outlier_years)

# Print the number of outlier years
print(num_outliers)
```
\vspace{1em}

5. Now calculate **Pearson's skew for the Yearly Mean Snow Depth Data**. Enter your
answer rounded to 2 decimal places. (do this in R; you've already proved your excel
skills above)   
```{r}
# Calculate Pearson's skew
skew <- skew(df_snow$Mean_Dec_May_snow_depth)
# Print the skew rounded to 2 decimal places
print(round(skew, 2))
```
\vspace{1em}

9. Calculate the **standard error of skew (ses)** for this data so we can determine the
significance of our skew. 
```{r}
# Calculate the skewness and standard error of skew
ses <- round(sqrt(6/length(df_snow$Mean_Dec_May_snow_depth)),2)
ses
```
\vspace{1em}

10. Based on the ses technique, is this **skew significant**?  
**No**
\vspace{1em}

11. Now determine the **kurtosis for the Yearly Mean Snow Depth data**. 
```{r}
kurtosis <- round(kurtosis(df_snow$Mean_Dec_May_snow_depth, type = 1) - 3, 2)
kurtosis
```
\vspace{1em}

12. Calculate the **standard error of kurtosis (sek)** for this data so we can determine the significance of our skew. 
```{r}
# Calculate standard error of kurtosis (SES)
ses_kurt <- round(sqrt(24/length(df_snow$Mean_Dec_May_snow_depth)),2)
ses_kurt
```
\vspace{1em}

13. Based on the sek technique, is this **kurtosis significant**? 
**No**
\vspace{1em}

14. Based on your statistical summary values, do you think that your data is normally
distributed?
**No**
\vspace{1em}

15. Is your data normally distributed? How can you tell? Be sure to report your certainty of this conclusion.
```{r}
shapiro.test(df_snow$Mean_Dec_May_snow_depth)
```
\vspace{1em}
 
 
```{r}
hist(df_snow$Mean_Dec_May_snow_depth)
```
 
